Proof-Reconstruction in Propositional Logic

### 2017-10-20 paper/sections/conclusions.tex
[ paper ] .0.

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proof-objects, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{sssec:normalization} and Section~\ref{sssec:reordering}.
We chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

@@

One of the main contribution of this study was
increasing the trustworthiness of the automatic provers,
in our case, \Metis.
Justifying a proof-object of a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.

>>>>

The reconstruction approach in this study was designed to use
only syntactical aspects of the logic.
This decision was in the beginning a drawback
since it demands more detailed proofs, a
description of every transformation or deduction step performed by
the prover, which is rarely included in the output of these programs,
see Section~\ref{sssec:normalization} and Section~\ref{sssec:reordering}.
We chose that syntactical treatment as the opportunity
to extend this work towards the support of first-order logic.
Recall, satisfiability in first-order logic is undecidable.

@@

One of the main contribution of this study was
increasing the trustworthiness of the automatic provers,
in our case, \Metis.
Justifying a proof of a theorem prover
has a real significant impact for these automatic tools.
The reverse engineering task can reveal important issues or bugs
in many parts of these systems (e.g., preprocessing, reasoning, or
deduction modules). During this research, we had the opportunity
to improve \Metis by reporting some bugs\footnote{Issues No.~2,
No.~4. and commit \name{8a3f11e} in the \Metis' official repository
\url{https://github.com/gilith/metis}}.
Fortunately, all these problems were fixed quickly by Joe Hurd in
20170810 release.

### 2017-10-20 paper/sections/discussion.tex
[ paper ] .0.

We noted in \TSTP derivations that \Metis does not take into account
some inference rules for the \Metis' official documentation; we
refer to the \Metis' logic kernel.
Besides the six \Metis' inference rules from
Fig.~\ref{fig:metis-inferences}, its proof-objects, the \TSTP
derivations, exhibit an entirely different set of rules.

>>>>

We noted in \TSTP derivations that \Metis does not take into account
some inference rules for the \Metis' official documentation; we
refer to the \Metis' logic kernel.
Besides the six \Metis' inference rules from
Fig.~\ref{fig:metis-inferences}, its proofs, the \TSTP
derivations, exhibit an entirely different set of rules.

### 2017-10-20 paper/sections/introduction.tex
[ paper ] .0.

Proof-reconstruction is a non-trivial task since it depends on the
integration between two complex systems.
We work in this study with one integration between an automatic theorem
prover (henceforth \ATP) and an interactive theorem prover (henceforth
\ITP).

The \ITP tools provide us the
logic framework to check and validate the reply of the \ATPs since
they allow us to define the formal language for the problems i.e.,
operators, logic variables, axioms, and theorems.

A proof-reconstruction tool provides such an integration, translating
the reply generated by the prover into the formalism of the proof
assistant.

% Since the formalism of the source (the evidence
% generated by the \ATP) is not necessarily the same logic in the
% target, the reconstruction turns out to be a ``reverse engineering''
% task.

Then, reconstructing a proof involves an in-depth
understanding of the algorithms in the \ATP and the logic specification
in the \ITP. To begin with the proof-reconstruction, it is
necessary to have a proof-object from the \ATP in a consistent
format, that is, a full script describing step-by-step
with exhaustive details and without ambiguities, the derivation to
get the actual proof.
For problems in classical propositional logic (henceforth \CPL),
from a list of at least forty\footnote{\ATPs available from the web
service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just a few
provers were able to deliver proofs (e.g.,
\name{CVC4}~\cite{Barrett2011}, \name{SPASS}, and
\name{Waldmeister}~\cite{hillenbrand1997}) and fewer replied with
a proof
in a file format like \TSTP~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}
(e.g., \name{E}, \Metis, \name{Vampire}, and \name{Z3}),
\len{LFSC}~\cite{Stump2008}, or the \len{SMT-LIB}~\cite{Bohme2011} format.

Many approaches have been proposed and some tools have been
implemented for proof-reconstruction in the last decades. These
programs are relevant not only because it helps to spread their
usage but they also increase the confidence of their users about
their algorithms and their correctness (see, for example, bugs in
\ATPs~\cite{Keller2013,Bohme2011,Fleury2014} and
\cite{Kanso2012}).
We mention some of these tools in Section~\ref{sec:related-work}.

In this paper, we ported a subset of the \Metis' inference rules to
\Agda, the propositional fragmented. The formalization of
these rules allowed us to justify step-by-step \Metis' \TSTP derivations by
translating automatically to \Agda by a proof-reconstruction tool built for
this research named \Athena.
This paper has been organized in the following way.
In Section~\ref{ssec:CPL}, we introduce some
notation and syntax of propositional logic.
In Section~\ref{sec:metis-language-and-proofs}, we introduce the
\Metis prover and its deduction system.
In Section~\ref{sec:proof-reconstruction}, we present our
approach to reconstruct proofs deliver by \Metis in type theory.
Section~\ref{sec:conclusions} summarizes the main findings and includes a
discussion of future work.
Some related works to this paper are presented in Section~\ref{sec:related-work}.
Finally, in appendix \textsc{A}, we provide a small example of the
proof-reconstruction workflow.

>>>>

% Proof-reconstruction is a non-trivial task since it depends on the
% integration between two complex systems.
% We describe one proof-reconstruction approach between an automatic theorem
% prover (henceforth \ATP) and an interactive theorem prover (henceforth
% \ITP).

% The \ITP tools provide us the
% logic framework to check and validate the reply of the \ATPs since
% they allow us to define the formal language for the problems i.e.,
% operators, logic variables, axioms, and theorems.

% A proof-reconstruction tool provides such an integration, translating
% the reply generated by the prover into the formalism of the proof
% assistant.

% % Since the formalism of the source (the evidence
% % generated by the \ATP) is not necessarily the same logic in the
% % target, the reconstruction turns out to be a ``reverse engineering''
% % task.

% Then, reconstructing a proof involves an in-depth
% understanding of the algorithms in the \ATP and the logic specification
% in the \ITP. To begin with the proof-reconstruction, it is
% necessary to have a proof from the \ATP in a consistent
% format, that is, a full script describing step-by-step
% with exhaustive details and without ambiguities, the derivation to
% get the actual proof.

% For problems in classical propositional logic (henceforth \CPL),
% from a list of at least forty\footnote{\ATPs available from the web
% service \name{SystemOnTPTP} of the TPTP World.} \ATPs, just a few
% provers were able to deliver proofs (e.g.,
% \name{CVC4}~\cite{Barrett2011}, \name{SPASS}, and
% \name{Waldmeister}~\cite{hillenbrand1997}) and fewer replied with
% a proof
% in a file format like \TSTP~\cite{Sutcliffe-Schulz-Claessen-VanGelder-2006}
% (e.g., \name{E}, \Metis, \name{Vampire}, and \name{Z3}),
% \len{LFSC}~\cite{Stump2008}, or the \len{SMT-LIB}~\cite{Bohme2011} format.

% Many approaches have been proposed and some tools have been
% implemented for proof-reconstruction in the last decades. These
% programs are relevant not only because it helps to spread their
% usage but they also increase the confidence of their users about
% their algorithms and their correctness (see, for example, bugs in
% \ATPs~\cite{Keller2013,Bohme2011,Fleury2014} and
% \cite{Kanso2012}).
% We mention some of these tools in Section~\ref{sec:related-work}.

% In this paper, we ported a subset of the \Metis' inference rules to
% \Agda, the propositional fragmented. The formalization of
% these rules allowed us to justify step-by-step \Metis' \TSTP derivations by
% translating automatically to \Agda by a proof-reconstruction tool built for
% this research named \Athena.
% This paper has been organized in the following way.
% In Section~\ref{ssec:CPL}, we introduce some
% notation and syntax of propositional logic.
% In Section~\ref{sec:metis-language-and-proofs}, we introduce the
% \Metis prover and its deduction system.
% In Section~\ref{sec:proof-reconstruction}, we present our
% approach to reconstruct proofs deliver by \Metis in type theory.
% Section~\ref{sec:conclusions} summarizes the main findings and includes a
% discussion of future work.
% Some related works to this paper are presented in Section~\ref{sec:related-work}.
% Finally, in appendix \textsc{A}, we provide a small example of the
% proof-reconstruction workflow.

### 2017-10-20 paper/sections/metis-language-and-proofs.tex
[ paper ] diagram: some fixes.
[ paper ] .0.

\subsection{Metis: Language and Proofs}
\label{ssec:metis-language-and-proofs}

\Metis is an automatic theorem prover written in \len{ML}
for first-order logic with equality developed by John Hurd
\cite{hurd2003first}. This prover is suitable for proof
reconstruction since it provides well-documented proof-objects
to justify its deduction steps from the basis of only six inference
rules.
In Fig.~\ref{fig:metis-inferences}, we show the three \Metis rules for
propositional logic.

@@

\subsubsection{Input Language.}
\label{ssec:input-language}

The \TPTP language --which includes the first-order
form (denoted by \name{fof}) and clause normal form (denoted by
\name{cnf}) formats~\cite{sutcliffe2009,Sicard-Ramirez2016} -- is de facto input
standard language to encode problems for many \ATPs. The \TPTP
syntax describes a well-defined grammar to handle annotated formulas
with the following form.

@@

The \name{language} can be \name{fof} or \name{cnf}. The \name{name}
serves to identify the formula within the problem. Each formula
assumes one \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, or a \name{hypothesis}.
The formulas include Boolean connectives (\verb!&!, \verb!|!,
\verb!=>!, \verb!<=>!, \verb!~!) and logic constants \verb!$true!
and \verb!$false!, for the symbols (∧, ∨, ⇒, ⇔, ¬, ⊤, ⊥), respectively.

@@

\subsubsection{Output Language.}
\label{ssec:output-language}

@@

The \name{rule} stands for the inference name; the other fields are
supporting arguments or useful information to apply the reasoning
step, and the list of parents nodes. For instance, in Fig.
\ref{fig:metis-proof-tstp}, \strip for example is the name of one of the inference with no arguments and only one parent node, the
\name{goal}.

@@

\subsubsection{Proof Rules.}
\label{sssec:proof-rules}

Since the scope of the problems treated in this research is \CPL, from the
list of \Metis inference rules in Fig.~\ref{fig:metis-inferences}, we found
that only three rules apply: \emph{axiom}, \emph{assume} and \emph{resolve}
rule.
Nonetheless, \TSTP derivations replied by \Metis use other rules. They are
\canonicalize, \clausify, \conjunct, \simplify, \negate, and \strip.

\Metis looks for a refutation to prove \verb!fof! problems.
A \verb!fof! problem contains the conjecture, the \emph{goal}, in the
problem. This goal is stripped into subgoals by using the \strip rule. The
subgoals are the conjuncts in a general conjunction that is equivalent to the
aforementioned goal.
Therefore, to prove the goal, \Metis proves independently each subgoal since
they tend to be less complex or even fewer propositions. At the end, the
conjunction of all subgoal proofs must prove the goal. Nevertheless, this last
reasoning is not explicit neither in the \TSTP derivations nor in the
documentation available about \Metis.
Since \Metis replies refutation proofs, to prove each subgoal, in the \TSTP
derivation, the inference rule \negate introduces the negation of such
subgoals. The \negate rule is a specification of the \emph{assume} rule.

\subsubsection{Proofs.}
\label{sssec:metis-proofs}

>>>>

\section{Metis: Language and Proofs}
\label{sec:metis-language-and-proofs}

\Metis is an automatic theorem prover written in \len{ML}
for first-order logic with equality developed by John Hurd
\cite{hurd2003first}. This prover is suitable for proof
reconstruction since it provides well-documented proofs
to justify its deduction steps from the basis of only six inference
rules.
In Fig.~\ref{fig:metis-inferences}, we show the three \Metis rules for
propositional logic.

@@

\subsection{Input Language.}
\label{ssec:input-language}

The \TPTP language --which includes the first-order
form (denoted by \name{fof}) and clause normal form (denoted by
\name{cnf}) formats~\cite{sutcliffe2009,Sicard-Ramirez2016} -- is the
input language to encode problems used by \Metis.
The \TPTP syntax describes a well-defined grammar to handle annotated
formulas with the following form.

@@

The \name{language} can be \name{fof} or \name{cnf}. The \name{name}
serves to identify the formula within the problem. Each formula
assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, or a \name{hypothesis}.
%existen otros
%----Keywords:    fof cnf thf tff include
%----Punctuation: ( ) , . [ ] :
%----Operators:   ! ? ~ & | <=> => <= <~> ~| ~& * +
%----Predicates:  = != $true $false

%----Top of Page---------------------------------------------------------------
%----CNF formulae (variables implicitly universally quantified)
% <cnf_formula>          ::= <disjunction> | (<disjunction>)
% <disjunction>          ::= <literal> | <disjunction> <vline> <literal>
% <literal>              ::= <fof_atomic_formula> | ~ <fof_atomic_formula> |
%                            <fof_infix_unary>
The formulas include the binary Boolean connectives
(\verb!&!, \verb!|!, \verb!=>!, \verb!<=>!, \verb!~!) and logic constants \verb!$true!
% the boolean type ($true and $false)
and \verb!$false!, for the symbols (∧, ∨, ⇒, ⇔, ¬, ⊤, ⊥), respectively.

@@

\subsection{Output Language.}
\label{ssec:output-language}

@@

The \name{rule} stands for the inference name; the other fields are
supporting arguments or useful information to apply the reasoning
step, and the list of parents nodes.

\begin{verbatim}
fof(subgoal_0, plain, p, inference(strip, [], [goal])).
\end{verbatim}

In the script above, \strip is for example the name of the inference.
It has no arguments and derives from one parent node with goal name.
\name{goal}.

\subsection{Inference Rules.}
\label{ssec:metis-inferences-rules}

% \TSTP derivations replied by \Metis use other rules. They are
% \canonicalize, \clausify, \conjunct, \simplify, \negate, and \strip.

% \Metis looks for a refutation to prove \verb!fof! problems.
% A \verb!fof! problem contains the conjecture, the \emph{goal}, in the
% problem. This goal is stripped into subgoals by using the \strip rule. The
% subgoals are the conjuncts in a general conjunction that is equivalent to the
% aforementioned goal.
% Therefore, to prove the goal, \Metis proves independently each subgoal since
% they tend to be less complex or even fewer propositions. At the end, the
% conjunction of all subgoal proofs must prove the goal. Nevertheless, this last
% reasoning is not explicit neither in the \TSTP derivations nor in the
% documentation available about \Metis.
% Since \Metis replies refutation proofs, to prove each subgoal, in the \TSTP
% derivation, the inference rule \negate introduces the negation of such
% subgoals. The \negate rule is a specification of the \emph{assume} rule.

\subsection{Proofs.}
\label{ssec:metis-proofs}

@@

### 2017-10-20 paper/sections/proof-reconstruction.tex
[ paper ] .0.

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct.
In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user, checking for the error in \Metis, \Athena,
or in \Agda.

@@

\begin{definition}[build$_{∨}$]
\begin{equation*}
  \begin{split}
  \label{eq:build-or}
  &\rm{build}_{∨} : \Prop → \Prop → \Prop\\
  &\rm{build}_{∨}(φ, ψ) =%\\
  % &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } \rm{build}_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation*}
\end{definition}

@@

\begin{lemma}[lem-build$_{∨}$]
\label{lem:lem-build-or}
$Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}(φ)$.
\end{lemma}

@@

\begin{definition}[sbuild$_{∨}$]
\begin{align*}
  \begin{split}
  \label{eq:strong-build-or-definition}
    &\rm{sbuild}_{∨} : \Prop → \Prop → \Prop\\
    &\rm{sbuild}_{∨}(φ, ψ) =%\\
    % &\hspace{3mm}
    \begin{cases}
      \ffactor(\rm{build}_{∨}(φ₁, ψ) ∨ \rm{build}_{∨}(φ₂,ψ)),
      &\text{ if }φ≡ φ₁ ∨ φ₂\\
      \rm{build}_{∨}(φ, ψ),  & \text{ otherwise.}
    \end{cases}
  \end{split}
\end{align*}
\end{definition}

\begin{lemma}[lem-sbuild$_{∨}$]
\label{lem:lem-sbuild-or}
$Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{definition}[reorder$_{∨}$]
\begin{align*}
    \label{eq:reorder-or}
    \begin{split}
    &\rm{reorder}_{∨} : \Prop \to \Prop \to \Prop\\
    &\rm{reorder}_{∨} (φ, ψ) = \rm{sbuild}_{∨}(rdisj(φ), ψ).
    \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-reorder$_{∨}$]
  \label{thm-reorder-or}
  $Γ ⊢ φ$, and $ψ: \Prop$ then $Γ ⊢ \rm{reorder}_{∨}(φ, ψ)$.
\end{theorem}

@@

\begin{definition}[reorder$_{∧}$]
  \begin{align*}
      \begin{split}
        &\rm{reorder}_{∧} : \Prop → \Prop \to \Prop\\
        &\rm{reorder}_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{ if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\text{, }\rm{reorder}_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }\rm{reorder}_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\\
          \rm{conjunct}(φ, ψ), &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-reorder$_{∧}$]
  \label{lem:lem-reorder-and}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \rm{reorder}_{∧}(φ, ψ)$.
\end{lemma}

\begin{definition}[conjunct$_{∨}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &\fconjunct_{∨} : \Prop → \Prop \to \Prop\\
    &\fconjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{ if }φ ≡ ψ\\
      ψ, &\text{ if }\freorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂,\ \fconjunct_{∨}(φ, ψ₁) ≡ ψ₁,\\
        &\text{and } \freorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ \fconjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ \fconjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-conjunct$_{∨}$]
  \label{lem:lem-conjunct-or}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{lemma}

\begin{definition}[reorder$_{∧∨}$]
 \begin{align*}
      \begin{split}
      &\freorder : \Prop \to \Prop \to \Prop\\
      &\freorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{ if } φ≡ψ\\
          ψ, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂,\ \freorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }\freorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂\\
          \fconjunct_{∨}(φ, ψ) &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

\begin{theorem}[thm-reorder$_{∧∨}$]
  \label{thm:thm-reorder-and-or}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$.
\end{theorem}

@@

\begin{definition}[\abbre{NProp}]
  \NProp is the type for a formula in normalized negative normal form
  in which negations appear only in the literals and the expression is
  in terms only of ($⊥$, $⊤$, $¬$, $∧$, $∨$, $⊻$) connectives.
  The $⊻$ connective is the exclusive disjunction.
  We define the \NProp type in a similar way as we did for \Prop type.
\end{definition}

>>>>

Finally, we type-check the \Agda proof-term. If the type-checking
success, the \TSTP derivation delivered by \Metis is correct
module \Agda and the formalizations of \Metis inferences.

In that case \Agda outputs an interface file.
Otherwise, when type-checking fails, the failure must be
investigated by the user, checking for the error in \Metis, \Athena,
or in \Agda.

@@

\begin{definition}[build$_{\boldsymbol{∨}}$]
\begin{equation*}
  \begin{split}
  \label{eq:build-or}
  &\rm{build}_{∨} : \Prop → \Prop → \Prop\\
  &\rm{build}_{∨}(φ, ψ) =%\\
  % &\hspace{3mm}
  \begin{cases}
  ψ, &\text{ if } φ ≡ ψ\\
  ψ, &\text{ if } ψ ≡ ψ₁ ∨ ψ₂\text{, } \rm{build}_{∨}(φ, ψ_{i}) ≡ ψ_{i}\text{ for some }i=1, 2\\
  φ, &\text{ otherwise.}
  \end{cases}
  \end{split}
\end{equation*}
\end{definition}

@@

\begin{lemma}[lem-build$_{\boldsymbol{∨}}$]
\label{lem:lem-build-or}
$Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}(φ)$.
\end{lemma}

@@

\begin{definition}[sbuild$_{\boldsymbol{∨}}$]
\begin{align*}
  \begin{split}
  \label{eq:strong-build-or-definition}
    &\rm{sbuild}_{∨} : \Prop → \Prop → \Prop\\
    &\rm{sbuild}_{∨}(φ, ψ) =%\\
    % &\hspace{3mm}
    \begin{cases}
      \ffactor(\rm{build}_{∨}(φ₁, ψ) ∨ \rm{build}_{∨}(φ₂,ψ)),
      &\text{ if }φ≡ φ₁ ∨ φ₂\\
      \rm{build}_{∨}(φ, ψ),  & \text{ otherwise.}
    \end{cases}
  \end{split}
\end{align*}
\end{definition}

\begin{lemma}[lem-sbuild$_{\boldsymbol{∨}}$]
\label{lem:lem-sbuild-or}
$Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{definition}[reorder$_{\boldsymbol{∨}}$]
\begin{align*}
    \label{eq:reorder-or}
    \begin{split}
    &\rm{reorder}_{∨} : \Prop \to \Prop \to \Prop\\
    &\rm{reorder}_{∨} (φ, ψ) = \rm{sbuild}_{∨}(rdisj(φ), ψ).
    \end{split}
  \end{align*}
\end{definition}

\begin{theorem}[thm-reorder$_{\boldsymbol{∨}}$]
  \label{thm-reorder-or}
  $Γ ⊢ φ$, and $ψ: \Prop$ then $Γ ⊢ \rm{reorder}_{∨}(φ, ψ)$.
\end{theorem}

@@

\begin{definition}[reorder$_{\boldsymbol{∧}}$]
  \begin{align*}
      \begin{split}
        &\rm{reorder}_{∧} : \Prop → \Prop \to \Prop\\
        &\rm{reorder}_{∧}(φ, ψ) =
        \begin{cases}
          φ, &\text{ if }φ ≡ ψ\\
          ψ₁ ∧ ψ₂, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\text{, }\rm{reorder}_{∧}(φ, ψ₁) ≡ ψ₁\\
                  &\text{and }\rm{reorder}_{∧}(φ, ψ₂)≡ ψ₂\\
          φ,       &\text{ if } ψ ≡ ψ₁ ∧ ψ₂\\
          \rm{conjunct}(φ, ψ), &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-reorder$_{\boldsymbol{∧}}$]
  \label{lem:lem-reorder-and}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \rm{reorder}_{∧}(φ, ψ)$.
\end{lemma}

\begin{definition}[conjunct$_{\boldsymbol{∨}}$]
  \label{eq:conjunct-or-definition}
\begin{align*}
  \begin{split}
    &\fconjunct_{∨} : \Prop → \Prop \to \Prop\\
    &\fconjunct_{∨}(φ, ψ) =
    \begin{cases}
      ψ, &\text{ if }φ ≡ ψ\\
      ψ, &\text{ if }\freorder_{∨}(φ, ψ) ≡ ψ\\
      ψ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂,\ \fconjunct_{∨}(φ, ψ₁) ≡ ψ₁,\\
        &\text{and } \freorder_{∨}(φ, ψ₂) ≡ ψ₂\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ \fconjunct_{∨}(φ₁, ψ) ≡ ψ\\
      ψ, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ \fconjunct_{∨}(φ₂, ψ) ≡ ψ\\
      φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-conjunct$_{\boldsymbol{∨}}$]
  \label{lem:lem-conjunct-or}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{lemma}

\begin{definition}[reorder$_{\boldsymbol{∧∨}}$]
 \begin{align*}
      \begin{split}
      &\freorder : \Prop \to \Prop \to \Prop\\
      &\freorder_{∧∨}(φ, ψ)=
        \begin{cases}
          ψ, &\text{ if } φ≡ψ\\
          ψ, &\text{ if } ψ ≡ ψ₁ ∧ ψ₂,\ \freorder_{∧∨}(φ,ψ₁) ≡ ψ₁\\
             &\text{and }\freorder_{∧∨}(φ, ψ₂) ≡ ψ₂\\
          φ, &\text{ if }ψ ≡ ψ₁ ∧ ψ₂\\
          \fconjunct_{∨}(φ, ψ) &\text{ otherwise.}
        \end{cases}
      \end{split}
  \end{align*}
  \end{definition}

\begin{theorem}[thm-reorder$_{\boldsymbol{∧∨}}$]
  \label{thm:thm-reorder-and-or}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$.
\end{theorem}

@@

\begin{definition}[\abbre{NProp}]
  \NProp is the type for a formula in normalized negative normal form
  in which negations appear only in the literals and the expression is
  in terms only of ($⊥$, $⊤$, $¬$, $∧$, $∨$, $\overline{\vee}$) connectives.
  The $⊻$ connective is the exclusive disjunction.
  We define the \NProp type in a similar way as we did for \Prop type.
\end{definition}
